Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-19 paper/paper.tex
more

\subsection{The free-energy principle and active inference \textbf{XXX move to Appendix}}
% \subsubsection{The theory}
In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
as biomechanical inference engine which must minimize the long-term average of surprise.
Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}. According to FEP's account, the goal of the brain is then to optimize over the generative model $G$: to iteratively
modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles. These is called the \textit{generative} model. FEP also postulates a dual model called the \textit{recognition} model,
which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams / fantasizes imaginary worlds $\z$ whilst the latter tries to generate sensations $\x$ which match these dreams!

@@

% \begin{figure}[!tb]
%   \centering
%   \includegraphics[width=.8\linewidth]{vae.pdf}
%   \caption{Variational autoencoders...}
%   \label{fig:vae}
% \end{figure}
The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.

>>>>

\subsection{The free-energy principle and active inference}
% \subsubsection{The theory}
In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
as biomechanical inference engine which must minimize the long-term average of surprise.
Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}. According to FEP's account, the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles. These is called the \textit{generative} model. FEP also postulates a dual model called the \textit{recognition} model,
which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams / fantasizes imaginary worlds $\z$ whilst the latter tries to generate sensations $\x$ which match these dreams!

@@

% \begin{figure}[!tb]
%   \centering
%   \includegraphics[width=.8\linewidth]{vae.pdf}
%   \caption{Variational autoencoders...}
%   \label{fig:vae}
% \end{figure}
The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
for minimizing free-energy, VAEs can be efficiently trained via back-propagation of prediction errors.
